---
title: 'CodeGrade: Intro'
author: "HulkMeyer"
date: "`r Sys.Date()`"
output: pdf_document
---

# CodeGrade Creation
Submit an R script file

Write programs in Markdown file and move code to R Script file. 

When running your code
  Loading packages use suppress to stop the loading packages output.  It will throw off CodeGrade
          suppressPackageStartupMessages("""put library(package)here""")
  Similarly, when you load CSVs do the same thing for same reason
          suppressMessages('''read_csv(csv file name)""")
          
Recommended psych package to gain access to the describe function
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(psych))
      #contains the describe function - for looking at your data
```
Use the describe function and what it gives you
```{r}
data("sleep")
summary(sleep)
```
```{r}
describe(sleep)
```
There is also a describe in Hmisc that looks very different
```{r warning=FALSE}
library(Hmisc)
describe(sleep)
```
If loaded both psych and Hmisc and  you want to use one or the other describe you would reference the package and then the function like so
package::fuction
```{r}
psych::describe(sleep)
```
Pizza Assignment - Don't have anything print automatically CodeGrade will command what it wants printed

Q1 <- pizza %>%
  group_by(date) %>%
  summarise(mean_pizza = mean(pizzas), .groups = 'drops') %>%
  arrange(desc(mean_pizza)) %>%
  select(date) %>%
  head(1)
  
Q2 <- pizza %>%
  mutate(hot = ifelse(temperature >= 72, 'Warm', 'Cool')) %V%
  group_by(hot) %>%
  summarise(mean_bill = mean(bill, na.rm=TRUE), .group = 'drop')

Q3 <- pizza %>%
  group_by(free_wine == 1) %>%
  summarise(sd_bill = sd(bill, na.rm = TRUE), .groups = 'drop') %>%
  arrange(desc(sd_bill))
  
Q4 <- pizza %>%
  mutate(ratio = bill/pizza) %>%
  group_by(operator, branch) %>%
  summarise(median_ratio = median(ratio, narm = TRUE),
    var_ratio = var(ratio, na.rm=TRUE), .groups = 'drop') %>%
  arrange(median_ratio)
  
Q5 <- pizza %>%
  filter(got_wine == 1, temerature < 70) %>%
  group_by(driver, branch, operator) %>%
  summarise(var_bill = var(bill, na.rm = TRUE),
    n=n(), .groups = 'drop') %>%
  filter(n >= 10) %>%
  arrange(desc(var_bill)) %>%
  select(driver, branch, operator) %>%
  head(1)
  
# A Note about Rscript from the command line

# CodeGrade Example

# CodeGrade Errors

















