---
title: "Module 3 - More R Intro"
author: "HulkMeyer"
date: "`r Sys.Date()`"
output: pdf_document
---

# More R Intro

# Basics
* This opens up a new window with work flow and such. 

# Data Import
library(tidyverse)
## contains readr package

# Data Transformation
library(nycflights)

```{r}
library(tidyverse)
library(nycflights13)
flights
```
```{r}
filter(flights, month ==1, day == 1)
```
# Filter

jan1 <- filter(flights, month==1, day==1)
```{r}
jan1 <- filter(flights, month==1, day==1)
jan1
```
```{r}
(dec25 <- filter(flights, month==12, day==25))
#wrapping whole statement in () will print out so it doesn't need to be called specifically
```
#Comparisons

When conducting operations where the difference is extreamely small (think 40th decimal place) use the near function
```{r}
sqrt(2)^2 == 2
```
```{r}
near(sqrt(2)^2, 2)
```
or see case
```{r}
1/49 * 49 == 1
```
```{r}
near(1/49*49,1)
```
& <- and function
| <- or function
! <- not function

# Missing Values

If you want to determine if a value is missing, use -> is.na()
```{r}
head(is.na(flights))
```
filter() only includes rows where the condition is TRUE; it excludes boaht FALSE and NA values.  If you want to preserve missing valuse, as for them explicityly:
-------
df <- tribble(x = c(1,NA,3))
filter(df,x>1)
filter(df, is.na(x) | x>1)

# Textbook Exercises 5.2
## Find all flights that

### Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```

### Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest =='IAH' | dest == 'HOU')
```

### Were operated by United, American, or Delta
```{r}
filter(flights, carrier =='AA' | carrier =='DL' | carrier == 'UA')
```

### Departed in summer (July, August, and September)
```{r}
filter(flights, month >= 7, month <= 9)
```

### Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, arr_delay >120, dep_delay<=0)
```

### Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay >30)
```

### Departed between midnight and 6am (inclusive) the previous challenges?
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)
```


### How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
# from problem (Departed in summer (July, August, and September))
# can use the between operator to achieve the same in less code
filter(flights, between(month, 7,9))
```

### Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule?(NA * 0 is a tricky counterexample!)
```{r}
filter(flights, is.na(dep_time))
```
If the flight never left that could mean flights that were canceled. 

# Arrange
arrange(data sheet, list columns in order you want them)
```{r}
arrange(flights, year, month, day)
```
arrange(flights, desc(dep_delay))
#modification and arranging the data frame by the values in a column
```{r}
arrange(flights, desc(dep_delay))
```
# Textbook Exercises 5.3
How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_time)))
```

Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))
```

Sort flights to find the fastest (highest speed) flights.
```{r}
arrange(flights,desc(distance/air_time))
```

Which flights traveled the farthest? Which travelled the shortest?

```{r}
arrange(flights, desc(distance))
```
```{r}
arrange(flights, distance)
```

# Select
select() is not terribly useful with the flights data because we only have 19 variable, but you can see how it would be useful if we had hundreds or so variables
```{r}
#slelect columns by name
select(flights, year, month, day)
#select all columns between year and day
select(flights, year:day)
#select all columns except those from year to day
select(flights, -(year:day))
```
some of other functions
starts_with('abc')
ends_with('123')
contains('asdf')
matches('(.)\\1')
num_range('x',1:3)'
rename(flights, tail_num = tailnum)
select(flights, time_hour, air_time, everything())

# Textbook Exercises 5.4
Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}

```

What happens if you include the name of a variable multiple times in a select() call?
```{r}

```

What does the any_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
?one_of
```


Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME"))

```

# Mutate

mutate() = create new variables based on old variables
```{r}
mutate(flights, 
       gain = dep_delay - arr_delay,
       speed = distance / air_time)
```
you can use variable created in a mutate to create other variables within the same mutate
```{r}
mutate(flights, 
        gain = dep_delay - arr_delay,
        hours = air_time / 60,
        gain_per_hour = gain / hours)
```
you can use transmute()
this will create a df that only has the variables that you created. 
```{r}
transmute(flights, 
        gain = dep_delay - arr_delay,
        hours = air_time / 60,
        gain_per_hour = gain / hours)
```

# Summarise

It collapses a data frame into a single row
```{r}
summarise(flights, delay=mean(dep_delay, na.rm=TRUE))
```
most useful when used with other functions like group_by
```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay=mean(dep_delay, na.rm=TRUE))
```
This can be used with multiple things
```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay=mean(dep_delay, na.rm=TRUE), sd_delay=sd(dep_delay, na.rm = TRUE))
```
# Pipes

%>%  pipe

magrittr package that has pipes
  this is in tidyverse
### Pipes Example - chapter 18
```{r}
foo_foo %>%
  hop(thorugh = forest) %>%
  scoop(up = field_mice) %>%
  bop(on = head)
```

# Missing Values Part 2
```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(mean = mean(dep_delay))
```
none of the mean values computed because all the things have missing values.  So we will have to affirmatively do something about htat. 
```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(mean = mean(dep_delay, na.rm=TRUE))
```
```{r}
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled <- flights %>%
  group_by(year, month, day) %>%
  summarise(mean = mean(dep_delay))

not_cancelled
```

# Count

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarise(delay = mean(arr_delay))

ggplot(data = delays, mapping = aes(x=delays)) +
  geom_freqpoly(binwidth = 10)
```
```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarise( delay = mean(arr_delay, na.rm=TRUE),
             n = n())

ggplot(data = delays, mappying = aes(x = n, y = delays)) +
  geom_point(alpha = 1/10)
```
```{r}
delays %>%
  filter(n>25) %>%
  ggplot(mapping = aes(x=n, y=delays)) +
  geom_point(alpha=1/10)
```
Changing the data set to Lehman's batting

```{r}
batting <- as_tibble(Lahman::Batting)

batters <- batting %>%
  group_by(playerID) %>%
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )

batters %>%
  filter(ab > 100) %>%
  ggplot(mapping = aes(x=ab, y=ba)) +
  geom_point()+
  geom_smooth(se = FALSE)
```

```{r}
batters %>%
  arrange(desc(ba))
```
because of small sample sizes you can see people with perfect averages but that is due to being 1 for 1. or the inverse is also true.  
to avoid this throwing off your visuals or data, you would implement a count standard to put a minumum number of at bats in this case. 

# STAR Wars Example
```{r}
library(tidyverse)
data("starwars")
head(starwars)
```
```{r}
starwars %>%
  select(height, mass, species)
```
Get height and mass by species
```{r}
starwars %>%
  select(height, mass, species) %>%
  group_by(species) %>%
  summarise(avg_height = mean(height, na.rm=TRUE)) %>%
  arrange(avg_height)
```
```{r}
starwars %>%
  select(height, mass, species) %>%
  group_by(species) %>%
  summarise(avg_height = mean(height, na.rm=TRUE)) %>%
  arrange(desc(avg_height))
```
```{r}
starwars %>%
  select(height, mass, species) %>%
  group_by(species) %>%
  summarise(avg_height = mean(height, na.rm=TRUE),
            avg_mass = mean(mass, na.rm=TRUE)) %>%
  arrange(desc(avg_height))
```
```{r}
starwars %>%
  select(height, mass, species) %>%
  group_by(species) %>%
  summarise(avg_height = mean(height, na.rm=TRUE),
            avg_mass = mean(mass, na.rm=TRUE),
            n = n()) %>%
  arrange(desc(avg_mass))
```
# Star Wars Part 2
```{r}
ggplot(filter(starwars)) +
  geom_point(aes(height, mass))

```
Remove the outlier
```{r}
ggplot(filter(starwars, species != 'Hutt')) +
  geom_point(aes(height, mass))
```
add a gender parameter
```{r}
ggplot(filter(starwars, species != 'Hutt')) +
  geom_point(aes(height, mass, color = gender))
```
changer gender to a facetwrap rather than color changes
```{r}
ggplot(filter(starwars, species != 'Hutt')) +
  geom_point(aes(height, mass)) +
  facet_wrap(~gender)
```

# Tidy Data
Tidy Data explained and its rules
  1) Each variable must have its own column
  2) Each observation must have its own row
  3) Each value must have its own cell

# Pivots
Why are those the rules these are the two main reasons why. 
1) Most people aren't familar with the priniciples of tidy data, and it's hard to derive them yourself unless you spend alot of time working with data. 
2) Data is often organised to facilitate some use other than analysis.  For example, data is often organised to make entry as easy as possible. 

# More R Wrap-Up
# Test Code
```{r}
starwars %>%
       summarise(av_mass = mean(mass, na.rm = TRUE))
```
```{r}
summarise(starwars, av_mass = mean(mass, na.rm = TRUE))
```











