---
title: "msleep Assignment Description"
author: "HulkMeyer"
date: "`r Sys.Date()`"
output: pdf_document
---
# msleep Assignment Description
For this assignment, you must name your R file nycflights13.R
* For all questions you should load tidyverse and nycflights13.  You should not need to use any other libraries. 
  * If the tidyverse package is not installed, you’ll need to do a one-time installation from the Console Window in RStudio like this:
    install.packages("tidyverse")
    **You cannot attempt to install packages in code that you submit to CodeGrade. **
  * If the nycflights13 package is not installed, you’ll need to do a one-time installation from the Console Window in RStudio like this:
    install.packages("nycflights13")
    **You cannot attempt to install packages in code that you submit to CodeGrade.** 
  * Load tidyverse with: 
    **suppressPackageStartupMessages(library(tidyverse))**
  * Load nycflights13 with:
    **suppressPackageStartupMessages(library(nycflights13))**
  * The actual data set is called flights.
    See the nycflights13 package page and chapter 5 from the textbook for more info.
* Round all float/dbl values to two decimal places.
  * If your rounding does not work the way you expect, convert the tibble to a dataframe by using as.data.frame()
* All statistics should be run with variables in the order I state
  * E.g., “Run a regression predicting mileage from mpg, make, and type” would be:
    **lm(mileage ~ mpg + make + type...)**
    
## Setting up packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(psych))
msleep
```
```{r}
psych::describe(msleep)
Hmisc::describe(msleep)
summary(msleep)
head(msleep)

```
## Problem 1
What is the variance in total sleep for carnivores and those whose conservation status is 1c?
  * You should be in a data frame with a value called var
```{r}
Q1 <- msleep %>%
  filter(vore == 'carni' & conservation == 'lc' ) %>%
  summarise(var = round(var(sleep_total),2)) %>%
  as.data.frame()
 
Q1
```
# Problem 2
What rodent(order Rodentia) has the highest total sleep / REM sleep ratio?
  * Your answer should be in a tibble assigned to Q2
```{r}
Q2 <- msleep %>%
  filter(order == 'Rodentia') %>%
  mutate(ratio = round(sleep_total/sleep_rem, 2)) %>%
  arrange(desc(ratio)) %>%
  select(name) %>%
  head(1)

Q2
```
## Problem 3
How many primates have a bodyweight/braingwt ratio greater than 100?
  * The answer should be assigned to Q3.
```{r}
Q3 <- msleep %>%
  filter(order == 'Primates') %>%
  mutate(ratio = bodywt/brainwt) %>%
  filter(ratio >= 100) %>%
  count(ratio >= 100)

Q3
```
## Problem 4
Create two new variables, mean of sleep total(named mean_sleep) and variance of sleep total(named var_sleep), grouped by conservation with missing values removed. 
  * The answer should be a dataframe assigned to Q4.
```{r}
Q4 <- msleep %>%
  group_by(conservation) %>%
  mutate(mean_sleep = round(mean(sleep_total,na.rm = TRUE), 2)) %>%
  mutate(var_sleep = round(var(sleep_total,na.rm = TRUE), 2)) %>%
  distinct(conservation, mean_sleep, var_sleep) %>%
  arrange(conservation) %>%
  head(6) %>%
  as.data.frame()
 
Q4
```
# Problem 5
Provide the name(s) of all the domesticated herbivores that sleep more than 12 hours. 
  * Your answer should be in a tibble assigned to Q5.
```{r}
Q5 <- msleep %>%
  filter(conservation == 'domesticated' & vore == 'herbi' & sleep_total >= 12) %>%
  select(name)
 
Q5
```

```{r}
head(msleep
     )
```


















