---
title: "nycflights13 Assignment Description and building"
author: "HulkMeyer"
date: "`r Sys.Date()`"
output: pdf_document
---
# nycflights13 Assignment Description

For this assignment, you must name your R file nycflights13.R
* For all questions you should load tidyverse and nycflights13.  You should not need to use any other libraries. 
  * If the tidyverse package is not installed, you’ll need to do a one-time installation from the Console Window in RStudio like this:
    install.packages("tidyverse")
    **You cannot attempt to install packages in code that you submit to CodeGrade. **
  * If the nycflights13 package is not installed, you’ll need to do a one-time installation from the Console Window in RStudio like this:
    install.packages("nycflights13")
    **You cannot attempt to install packages in code that you submit to CodeGrade.** 
  * Load tidyverse with: 
    **suppressPackageStartupMessages(library(tidyverse))**
  * Load nycflights13 with:
    **suppressPackageStartupMessages(library(nycflights13))**
  * The actual data set is called flights.
    See the nycflights13 package page and chapter 5 from the textbook for more info.
* Round all float/dbl values to two decimal places.
  * If your rounding does not work the way you expect, convert the tibble to a dataframe by using as.data.frame()
* All statistics should be run with variables in the order I state
  * E.g., “Run a regression predicting mileage from mpg, make, and type” would be:
    **lm(mileage ~ mpg + make + type...)**

**Before attempting to answer these questions, please review all CodeGrade information provided in the CodeGrade: Intro module. If you do not, you are likely to lose points. **
## Setting up packages
```{r warning=FALSE, include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(nycflights13))
suppressPackageStartupMessages(library(psych))
```
## Take a look at the data
```{r}
psych::describe(flights)
```
```{r}
head(flights)
```
```{r}
summary(flights)
```

## Problem 1
** What is the mean distance of flights for each of the carriers AA, EV, and FL?
    * This answer should e a dataframe assigned to Q1.
```{r}
Q1 <- flights %>%
  group_by(carrier) %>%
  filter(carrier == 'AA' | carrier == 'EV' | carrier == 'FL') %>%
  summarise(avg_dis = round(mean(distance),2)) %>%
  as.data.frame()
```
## Problem 2
For the month with the highest number of flights what is the value?
  * The answer should be assigned to Q2. 
```{r}
Q2 <- flights %>%
  group_by(month) %>%
  summarise(NumFlights = n()) %>%
  arrange(desc(NumFlights)) %>%
  head(1)
```
## Problem 3
Find the five shortest minimum distances, called min_disct, by origin/destination combination.
  * The answer should be a tibble assigned to Q3.
```{r}
Q3 <- flights %>%
  mutate(min_dist = distance) %>% #creates min_distance parameter
  select(origin, dest, min_dist) %>% #selects columns for output
  arrange(min_dist) %>% # arrange from shortest to longest
  distinct(origin, dest, min_dist) %>% #Give distinct routes
  head(5) # outputs first 5 rows
Q3
```
## Problem 4
What five days of the year had the highest mean distance when leaving JFK?  Sort in Decending order.
  * The answer should be a datafraem assigned to Q4.
```{r}
Q4 <- flights %>%
  filter(origin == 'JFK') %>%
  group_by(month, day) %>%
  mutate(mean_distance = round(mean(distance,na.rm = TRUE),2)) %>%
  select(month, day, mean_distance) %>%
  distinct(month, day, mean_distance) %>%
  arrange(desc(mean_distance)) %>%
  head(5) %>%
  as.data.frame()
Q4
```
## Problem 5
Calculate the maximum arrival delay for flights to Boston and Atlanta, separately.
  * The answer should be a tibble assigned to Q5
```{r}
Q5 <- flights %>%
  group_by(carrier) %>%
  filter(dest == 'ATL' | dest == 'BOS') %>%
  group_by(dest) %>%
  summarise(max(arr_delay, na.rm = TRUE))
  
  
  
  
  
  
  
  
Q5
```
```{r}
flights
```


