---
title: "pizza Assignment Description and Building"
author: "HulkMeyer"
date: "`r Sys.Date()`"
output: pdf_document
---
# pizza Assignment Description
* For this assignment, name you R file pizza.R
  * For all questions you should load tidyverse.  You should not need to use any other libraries.
  * Load tidyverse with:
    **suppressPackageStartupMessages(library(tidyverse))**
* Download the pizza.csv file from Brightspace and place it in the same folder/directory as your script file.  Then in RStudio, set your working Directory to your Source File Location:
* Load the pizza.csv file like this:
    **pizza <- read_csv("pizza.csv")**
* Round all float/dbl values to two decimal places.
  * If your rounding does not work the way you expect, convert the tibble to a dataframe b using as.data.frame()
* All statistics should be run with variables in the order I state
  * Run a regression predicting mileage from mpg, make and type would be:
    ** lm(mileage ~ mpg + make + type ...)**
* In each of these you must atleast use two dplyr functions.  you may use Google to look up how to do certain aspects. 


## Setting up packages and data set
```{r message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(psych))
pizza <- read_csv("pizza.csv")
```
## Taking a look at the data
```{r}
psych::describe(pizza)
summary(pizza)
head(pizza)
pizza
```
## Problem 1
Create a tibble containing driver names of instances where free_wine=1, discount_customer = 1, and the order contained more than 4 pizzas
  (There will be repeated names)
  * Assign the tibble to Q1.
  * The list portion of your answer (which CodeGrade will b looking at) should look something like this:
```{r}
Q1 <- pizza %>%
  filter(free_wine == 1 & discount_customer == 1 & pizzas > 4) %>%
  select(driver) %>%
  as.data.frame()

Q1
```
## Problem 2
Create a variable name ratio that is the ratio of bill to pizza, called ratio.  What is the mean of that value (call the value mean_ratio)?
* Your answer should be a 1x1 dataframe.  Assign this to Q2
```{r}
Q2 <- pizza %>%
  mutate(ratio = bill/pizzas) %>%
  summarise(mean_ratio = round(mean(ratio),2)) %>%
  as.data.frame()

Q2
```
## Problem 3
For each day of the week, what is the variance in pizzas?
* The created values should be called var_pizzas
* The answer should be a dataframe assigned to Q3
```{r}
Q3 <- pizza %>%
  group_by(day) %>%
  summarise(var_pizzas = round(var(pizzas),2)) %>%
  as.data.frame()

Q3
```
## Problem 4
Which operator had the higher average bill?
* The answer should be a tibble assigned to Q4
```{r}
Q4 <- pizza %>%
  group_by(operator) %>%
  summarise(avg_bill = round(var(bill),2)) %>%
  arrange(desc(avg_bill)) %>%
  select(operator) %>%
  head(1)
  

Q4
```
## Problem 5
What was the highest amount of free wine given by day/driver combination?
(For instance, Friday Bruno was 13, while Wednesday Salvator was 12)
* The asnwer should be a tibble assigned to Q5.
```{r}
Q5 <- pizza %>%
  select(day, driver, free_wine) %>%
  group_by(day, driver) %>%
  summarise(n = as.integer(sum(free_wine))) %>%
  arrange(desc(n)) %>%
  head(1)
  
Q5
```



































